Foreign Exchange Speculation: An event study

### 2017-11-30 Knitr/FinancialEconomics.tex
Updated type to article.
Updated and corrected.

\documentclass[preprint,12pt,authoryear]{elsarticle}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
\begin{document}
\maketitle
\begin{abstract}
Does speculation facilitate price discovery or instability?  If it is price discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction. This paper seeks to distinguish between these two characteristics by analysing those times when speculation in the foreign exchange market is most extreme.  A series of event studies are conducted on the extremes of speculative activity and speculative sentiment. If speculation is noise,  extreme sentiment and extreme positions should lead to overshooting and increase risk of subsequent reversals. The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.

@@

\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} emphasised the role that speculation played in providing liquidity. Speculators provide a service for a fee that is determined by the return on their activity.  Informed speculators will tend to drive out the uninformed noise-traders \citet{FriedmanPositive}.  Successful speculators would be those buying below intrinsic value and selling above that rate.  However, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p. 101]{Keynes1936}.

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices. \footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes.

The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.

@@

\citet[p.529]{BlackNoise} argued that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to exchange. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  This space can encourage informed traders to pay the cost of acquiring information.  \citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.

\section{Measuring speculation}
This study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.

If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%. See \citet{Global} for a technical review. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.  See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.

@@

The effect of extreme speculation on future returns is analysed with an event study.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.

@@

\section*{References}
\bibliography{../../myrefs}

>>>>

\documentclass{article}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{multirow}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
\title{Foreign Exchange Speculation: An event study}
\author{Rob Hayward}
\date{Dec 2017}
\begin{document}
\maketitle
\begin{abstract}
Does speculation facilitate price discovery or instability?  If it is price discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction. This paper seeks to distinguish between these two characteristics by analysing those times when speculation in the foreign exchange market is most extreme.  A series of event studies are conducted on the extremes of speculative activity and speculative sentiment. If speculation is noise,  extreme sentiment and extreme positions should lead to overshooting and increase risk of subsequent reversals. The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency of financial markets.

@@

\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} for example emphasised the role that speculation plays in facilitating financial market transactions. Speculators offer a service for a fee that is determined by the return on their activity.  While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.   Nonetheless, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p. 101]{Keynes1936} and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge.

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes.

The rest of the paper proceeds as follows:  Section Two presents the framework of informed and uninformed speculation; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.

@@

\citet[p.529]{BlackNoise} asserted that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to exchange. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  This space can encourage informed traders to pay the cost of acquiring information in the \citet{Grossman1980Impossibility} model.  There is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.

\section{Measuring speculation}
This study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.    If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%. See \citet{Global} for a technical review. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.  See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.

@@

The effect of extreme speculation on future returns is analysed with an event study.  The events here are the extremes of speculative sentiment and speculative positions.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.

@@

\bibliography{../../myrefs}
